### 2024-05-26 paper/paper.tex
Wrote related work section for neural speaker anonymization and evasion attacks

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[preprint]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}

\graphicspath{{./images}}

\title{Attacks on Neural Networks in a Lightweight Speech Anonymization Pipeline}

\author{Daan Brugmans \\
  Radboud University\\
  \texttt{daan.brugmans@ru.nl}
}

\begin{document}

\maketitle
% \begin{table}
%     \centering
%     \begin{tabular}{cols}

%     \end{tabular}
%     \caption{}
%     \label{tab:}
% \end{table}

% \begin{figure}
%     \centering
%     \includegraphics[width=0.9\textwidth]{}
%     \caption{}
%     \label{fig:}
% \end{figure}

\begin{abstract}

\end{abstract}

\section{Introduction}
As advancements in the field of Automatic Speech Recognition (ASR) have accelerated with the rise of modern End-to-End neural networks, the risks associated with using such models in ASR applications has become more evident.

Modern neural ASR models are capable of parsing and producing speech to a new level of authenticity: transformers are State-of-the-Art for ASR and word recognition, and the introduction of modern unsupervised deep neural network architectures, such as Generative Adversarial Networks (GANs) and Variational Autoencoders (VAEs), has allowed for more realistic, accurate, and easier generation of speech.
These modern speech generation models are capable of learning to reproduce a person's voice, and then generating new speech using the learned voice.
Such synthesized speeches are called \textit{deepfaked} speeches, or simply \textit{deepfakes}.

The presence and influence of deepfakes has become increasingly apparent in recent years: neurally synthesized audio and video of important persons are used to spread misinformation and manipulate.
One way to counteract the repercussions of deepfakes is the removal of the personalization in the learned, and thus reproduced, speech.
This is called \textit{Speaker Anonymization}.
In Speaker Anonymization, we aim to maintain the ASR quality of the audio, while applying changes that make the audio untraceable to a person's likeness.

Although modern Speaker Anonymization systems, often neural in nature, have been shown to be able to anonymize speech while maintaining ASR quality, they themselves can also be manipulated.
By attacking neural Speaker Anonymization systems, we may be able to circumvent the preventative measures they provide, and generate speech to a person's likeness regardless of their presence.
This paper will focus in that topic: attacking neural Speaker Anonymization systems.
Specifically, we will focus on the topic of adversarial attacks.
\textit{Adversarial attacks} on deep neural networks are attacks that alter the input that the network receives.
They are performed by applying a perturbation onto the input data with the aim to alter or manipulate the network's behavior.
These perturbations should meet two criteria as faithfully as possible: they should not be detectable by humans and they should perturb the original input as little as possible.

\section{Related Work}
\subsection{Neural Speaker Anonymization}
\citet{meyer2023anonymizing} showcase a successful realization of neural Speaker Anonymization using GANs.
Their Speaker Anonymization architecture includes the extraction of embeddings from speech, which are fed into a GAN.
This GAN first learns to generate embeddings sampled from a normal distribution.
After every sampling step, it then learns to calculate the distance between the embedding that it generated, and an embeddings extracted from speech it has been fed.
This training procedure teaches the GAN to generate embeddings that are similar to the speech embeddings.
Once training has been finished, embeddings based on real speech embeddings are sampled from the GAN until an embedding is generated whose cosine distance from the corresponding real speech embeddings is sufficiently large.
This embedding is fed to an existing speech synthesis model, which produces anonymized speech.

\citet{shihao2024adversarial} showcase another example of a neural Speaker Anonymization pipeline.
Their main contribution is the use of an \textit{adversarial attack} for anonymization purposes.
\citeauthor{shihao2024adversarial} use an adversarial attack called \textit{FGSM} that learns to apply perturbations on a VAE's latent space vector that represents a personalized utterance.
The perturbations caused by the FGSM attack alter the latent space vector in such a way that it is as far removed from the original speaker as possible, while still being recognizable.
When the vector is then fed to the VAE's decoder, it is unable to extract features from the perturbed vector that relate to the original speaker's speech characterizations.
The resulting decoded speech sample is thus anonymous.

\subsection{Attacks on Neural Networks}
The FGSM attack used by \citet{shihao2024adversarial} is an example of an adversarial attack.
Adversarial attacks are extensively described by \citet{xiaoyong2019adversarial}.
They provide an introduction to and an overview of adversarial attacks within the deep learning domain, and should provide the reader with plentiful knowledge on the topic.

We will focus on two types of adversarial attacks: \textit{Evasion Attacks} and \textit{Backdoor Attacks}.

\subsubsection{Evasion Attacks}
Evasion attacks are adversarial attacks that are used during model inference.
The aim of an evasion attack is to perturb an input in such a way that a fully trained model is fooled into behaving differently.
This behavior should fulfill the attacker's goals.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.45\textwidth]{fgsm.png}
    \caption{An example of an FGSM attack visualized \citep{goodfellow2015explaining}.}
    \label{fig:fgsm}
\end{figure}

\citet{goodfellow2015explaining} introduced FGSM, the \textit{Fast Gradient Sign Method}.
FGSM perturbs the input by exploiting a trained model's gradients.
When given the input $x$ and its label $y$, an FGSM attack calculates the gradient with respect to the input using the model's parameters $\theta$ and loss function $J(\theta, x, y)$.
The sign of this gradient is calculated and is added on top of the original input by a factor of $\epsilon$.
The FGSM attack can then be defined as such:
\begin{align*}
  x_t &= x + \epsilon \cdot \text{sign}(\nabla_x J(\theta, x, y))
\end{align*}
where $x_t$ is the perturbed input.
A visualization of this process can be found in figure \ref{fig:fgsm}.

FGSM attacks have some limitations.
One of these limitations is that an FGSM attack is a one-step approach: the gradient with respect to the input is calculated once, and is then used immediately with to corrective measures.
Although this makes FGSM attacks cheap, it also makes the attack difficult to perform optimally.
\citet{madry2018towards} propose an expanded version of the FGSM attack called the \textit{Projected Gradient Descent} (PGD) method.
The core principle of PGD is that it projects FGSM gradients back to a predefined max perturbation level; if an FGSM's perturbation is too big, PGD projects it back.
This makes PGD a multiple-step approach.

PGD's projection of an FGSM attack is performed by applying a clipping function $\text{clip}()$ on the FGSM attack.
The clipping function will clip gradients outside the limit of $x + S$, where $x$ is the original input data and $S$ is the predefined maximal Euclidean distance a perturbation is allowed to be from $x$.
Prior to clipping the FGSM perturbed input $x_t$, PGD will apply another FGSM attack on the model, but will now calculate the gradient with respect to the perturbed input $x_t$ instead of the original input $x$.
This additional FGSM attack is added onto $x_t$ by a factor of hyperparameter $\alpha$.
It is the result of this addition that is clipped by the clipping function.
This means that a PGD attack can be defined as follows:
\begin{align*}
  x_t &= x + \epsilon \cdot \text{sign}(\nabla_x J(\theta, x, y)) \\
  x_{t+1} &= \text{clip}_{x + S}(x_t + \alpha \cdot \text{sign}(\nabla_x J(\theta, x_t, y)))
\end{align*}
where $x_{t+1}$ is the final result of one round of PGD.
PGD can be performed for multiple rounds $T$ until $x_T$ is reached.

\subsubsection{Backdoor Attacks}
\citet{gu2019badnets} % The paper about BadNets. Keep it brief. Use it to visualize the backdooring process.

\citet{liu2018trojaning} % The paper about trojanning audio. Keep it brief. Use it to show that backdooring audio is also possible.

\citet{stefanos2022ultrasonic} % The paper about ultrasonic frequencies for backdooring. Keep it brief. Another example of audio backdoor.

\citet{stefanos2023jingleback} % Jingleback paper. Explain in more details since you will focus on it.

\subsection{Attacks on Neural Speech Systems}
\citet{neekhara2019universal} % Paper about adversarial attacks on ASR systems

\citet{kreuk2018fooling} % Paper about adversarial attacks on E2E speaker verification

\section{Method}
\citet{kai2022lightweight} % Work on lightweight speaker anonymization, idea for our method/pipeline.

\section{Experiment}

\section{Results}

\section{Discussion}

\section{Conclusion}

\bibliography{custom}

\appendix

\end{document}